---
title: "MSA8190-HW6"
author: "Kristen Biskobing"
date: "October 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
#setwd("H:/Desktop/School/MSA8190/homeworks")
library("pwr")
library("zoo")
```

## Problem 1
A confidence interval estimate is desired for the gain in a circuit on a semiconductor device. Assume that gain is normally distributed with standard deviation sigma = 20.
```{r }
sigma <- 20
```

###(a) Find a 95% CI for mu when n = 10 and x_bar = 1000. alpha<-0.05
```{r }
alpha <-0.05 
n <- 10
x_bar<-1000
za<-qnorm(alpha/2, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
ciu<-x_bar+(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, `r ciu`]

###(b) Find a 95% CI for mu when n = 25 and x_bar = 1000. alpha<-0.05
```{r }
alpha <-0.05 
n <- 25
x_bar<-1000
za<-qnorm(alpha/2, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
ciu<-x_bar+(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, `r ciu`]

###(c) Find a 99% CI for mu when n = 10 and x_bar = 1000. alpha<-0.01
```{r }
alpha <-0.01 
n <- 10
x_bar<-1000
za<-qnorm(alpha/2, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
ciu<-x_bar+(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, `r ciu`]


###(d) Find a 99% CI for mu when n = 25 and x_bar = 1000. alpha<-0.01
```{r }
alpha <-0.01
n <- 25
x_bar<-1000
za<-qnorm(alpha/2, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
ciu<-x_bar+(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, `r ciu`]

##Problem 2
A manufacturer produces piston rings for an automobile engine. It is known that ring
diameter is normally distributed with sigma = 0.001 millimeters. A random sample of 15 rings
has a mean diameter of x_bar = 74.036 millimeters.
```{r }
sigma <- 0.001
x_bar <- 74.036
n <- 15
```
###(a) Construct a 99% two-sided confidence interval on the mean piston ring diameter.
```{r }
alpha <- 0.01
za<-qnorm(alpha/2, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
ciu<-x_bar+(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, `r ciu`]

###(b) Construct a 95% lower-confidence bound on the mean piston ring diameter.
```{r }
alpha <- 0.05
za<-qnorm(alpha, lower.tail=FALSE)
cil<-x_bar-(za*(sigma/sqrt(n)))
```
The confidence interval is [`r cil`, \infinity ]

##Problem 3
A manufacturer produces crankshafts for an automobile engine. The wear of the crankshaft
after 100,000 miles (0.0001 inch) is of interest because it is likely to have an impact on
warranty claims. A random sample of n = 15 shafts is tested and x_bar = 2.78. It is known that
sigma = 0.9 and that wear is normally distributed.
```{r }
sigma <- 0.9
x_bar <- 2.78
n <- 15
```
###(a) Test H0 : mu = 3 versus H1 : mu != 3 using alpha = 0.05
```{r }
mu <- 3
alpha <- 0.05
zk<-(x_bar-mu)/(sigma/sqrt(n))
za<-qnorm(alpha/2, lower.tail=FALSE)
abs(zk)>za
```
So mu=3 is not rejected. 

###(b) What is the power of this test if mu = 3.25?
```{r }
d<-(3.25-2.78)/sigma
pwr.norm.test(d=d,sig.level=alpha, n=n )
```
###(c) What sample size would be required to detect a true mean of 3.75 if we wanted the power to be at least 0.9?
```{r }
d<-(3.75-2.78)/sigma
pwr.norm.test(d=d,sig.level=alpha, power=0.9)
```
###(d) Explain how the question in part (a) could be answered by using p-value and confidence interval.
The confidence interval could be found using the below code.
```{r }
  cil<-x_bar-(za*(sigma/sqrt(n)))
  ciu<-x_bar+(za*(sigma/sqrt(n)))
```
Since `r mu` is between `r cil` `r ciu` we do not reject the hypothesis. 

The p-value is calculated by taking pnorm of the computed test statistic and applying the following:
```{r }  
  p<-2*(1-pnorm(zk,0,1))
  p<alpha
```  
Since p<alpha is false, we do not reject the hypothesis. 
  
##  Problem 4
###Calculate the following probabilities.
###(a) If X1,X2,X3 ~ NID(mu = 1, sig2 = 1), what is P(X1 + X2 > X3)?
P(X1 + X2 > X3) = P(X1+X2-X3>0) 
E(X1+X2-X3) = 1+1-1=1
Var(X1+X2-X3) = 1+1-1=1
so X1+X2-X3~NID(1,1) and 
```{r }
pnorm(0,1,sqrt(1), lower.tail=FALSE)
```

###(b) If X1,X2 ~ NID(mu = 0, sig2), what is P(X_bar < S)?
P(X-bar<S) = P(X-bar-S<0) 
X_bar~N(0,S^2/n)
(n-1)S^2/sigma^2~X^2(n-1)
E(S^2)=sigma^2
E(X-bar-S) = 1+1-1=1
Var(X-bar-S) = 1+1-1=1
so X-bar-S~NID(1,1) and 
```{r }
pnorm(0,1,sqrt(1))
```

###(c) If X1,X2 ~ NID(mu = 0, sig2), what is P(X1 + X2 < sqrt(2(X1-X2)^2)?
P(X1 + X2 < sqrt(2(X1-X2)^2)) = P(X1+X2-sqrt(2(X1-X2)^2)<0) 
E(X1+X2-sqrt(2(X1-X2)^2)) = 0+0-0=0
Var(X1+X2-sqrt(2(X1-X2)^2)) = 1+1-1=1 ???? 
so X1+X2-sqrt(2(X1-X2)^2)~NID(0,???) and 
```{r }
pnorm(0,1,sqrt(1))
```

###(d) If X1,X2 ~ NID(mu = 0, sig2), Y1,Y2 ~ NID(mu = 0; 2sig2), and X's and Y 's are independent, what is
P(X1 + X2 > X3) = P(X1+X2-X3>0) 
E(X1+X2-X3) = 1+1-1=1
Var(X1+X2-X3) = 1+1-1=1
so X1+X2-X3~NID(1,1) and 
```{r }
pnorm(0,1,sqrt(1), lower.tail=FALSE)
```

##Problem 5
The compressive strength of concrete is being tested by a civil engineer. He tests 12 specimens and obtains the following data.
```{r } 
d5<-c(2216,2237,2225,2301,2318,2255,2249,2204,2281,2263,2275,2295)
n<-12
```
###(a) Is there evidence to support the assumption that compressive strength is normally distributed? Does this data set support your point of view? Include a graphical display in your answer.
```{r } 
x_bar<-mean(d5)
s<-sd(d5)

alpha<-0.05
qqnorm(d5); qqline(d5)
```
Yes, the data seens to follow a normal distribution. The probability plot is very accurate close to the 0 quantile and varies more the farther out you go.  

###(b) Test the normality of the data by Shapiro-Wilk test?
```{r } 
shapiro.test(d5)
```
The test gives a very high p value, much larger than any normal alpha, so we can not reject that this has a normal distribution.

###(c) Construct a 95% two-sided confidence interval on the mean strength.
```{r } 
ta<-qt(0.025, n-1, lower.tail=FALSE)
cil<-x_bar-(ta*(s/sqrt(n)))
ciu<-x_bar+(ta*(s/sqrt(n)))  
```
The confidence interval is [`r cil`, `r ciu`].

###(d) Construct a 95% lower-confidence bound on the mean strength.
```{r } 
ta<-qt(0.05, n-1, lower.tail=FALSE)
cil<-x_bar-(ta*(s/sqrt(n)))
```
The confidence interval is [`r cil`, \infinity].

##Problem 6
The sugar content of the syrup in canned peaches is normally distributed. Suppose that the
variance is thought to be sigma^2 = 18 (milligrams). A random sample of n = 10 cans yields a
sample standard deviation of s = 4.8 milligrams.  
```{r } 
var<-18
n<-10
s<-4.8
```
###(a) Test the hypothesis H0 : sigma^2 = 18 versus H1 : sigma^2 != 18 using alpha = 0.05.
Have to check both sides since not symmetric
```{r } 
alpha<-0.05
chi20<-(((n-1)*s^2)/var)
chia2l<-qchisq(alpha/2,n-1)
chia2u<-qchisq(alpha/2,n-1, lower.tail=FALSE)
chi20>chia2u
chi20<chia2l
```
Since it is not smaller than the lower bound or larger than the upper bound, we can not reject the null hypothesis.

###(b) What is the P-value for this test?
```{r } 
2*(1-pchisq(chi20, n-1))
```
###(c) Find a 95% two-sided confidence interval for sigma.
```{r } 
clb<-(((n-1)*s^2)/chia2u)
cub<-(((n-1)*s^2)/chia2l)
```
The confidence interval is [`r clb`, `r cub`].

###(d) Find a 90% lower confidence bound for sigma.
```{r } 
(((n-1)*s^2)/qchisq(0.1,n-1, lower.tail=FALSE))
```
###(e) Discuss how part (a) could be answered by constructing a 95% two-sided confidence interval for sigma.

The hypothesis could also be tested by conducting a confidence interval using the sample variance. If the expected value of the variance falls outside of the interval then the hypothesis is rejected. 

###(f) Suppose that the true variance is sigma2 = 40. How large a sample would be required to detect this difference with probability at least 0.90?
```{r }
lambda<-sqrt(40)/s
B<-0.90 
pwr.chisq.test(w=lambda, df=n-1, sig.level=alpha/2, power=B)
```
Using the pwr.chisq.test an N=12 would be needed. 

##Problem 7
The rainfall in acre-feet from 20 clouds that were selected at random and seeded with silver
nitrate follows:
```{r } 
d7<-c(18.0,30.7,19.8,27.1,22.3,18.8,31.8,23.4,21.2,27.9,31.9,27.1,25.0,24.7,26.9,21.8,29.2,34.8,26.7,31.6)
n<-20
x_bar<-mean(d7)
s<-sd(d7)
```
###(a) Can you support a claim that mean rainfall from seeded clouds exceeds 25 acre-feet? Use alpha = 0.01. 
H0:x=25
H1:x>25
```{r } 
alpha<-0.01
x0<-25
t0<-(x_bar-x0)/(s/sqrt(n))
ta<-qt(alpha, n-1)
t0>ta
```
Since t0>ta we reject the hypothesis that it is =25, so we can support the claim. 

###(b) Is there evidence that rainfall is normally distributed?
```{r } 
shapiro.test(d7)
qqnorm(d7); qqline(d7)
```
With alpha as 0.01 then we can't reject from the shapiro test and the probability plot shows a normal distribution fits relatively well. 

###(c) Compute the power of the test if the true mean rainfall is 27 acre-feet.
```{r } 
d<-(27-x_bar)/s
pwr.norm.test(d=d,sig.level=alpha, n=n)
```

###(d) What sample size would be required to detect a true mean rainfall of 27.5 acre-feet if we wanted the power of the test to be at least 0.9?
```{r } 
d<-(27.5-x_bar)/s
pwr.norm.test(d=d,sig.level=alpha, power=0.9)
```
n=158.7 so 159 samples

###(e) Explain how the question in part (a) could be answered by constructing a one-sided confidence bound on the mean diameter.
```{r } 
ta<-qt(alpha, n-1)
ci<-x_bar-(ta*(s/sqrt(n)))
x0>ci
```
since x0 does not fall within the confidence interval [`r ci`-\infinity), we reject the hypothesis that the mean = 25 and can then assume mean > 25.

###(f) Explain how the question in part (a) could be answered by using p-value.
u>25
```{r } 
p<-pt(t0, n-1, lower.tail=FALSE)
p<alpha
```  
Since p<alpha is false we can not reject the hypothesis that mean=25. 

##Problem 8
The sodium content of thirty 300-gram boxes of organic corn flakes was determined. The data (in milligrams) are as follows:
```{r } 
d8<-c(131.15,130.69,130.91,129.54,129.64,128.77,130.72,128.33,128.24,129.65,130.14,129.29,128.71,129.00,129.39,130.42,129.53,130.12,129.78,130.92,131.15,130.69,130.91,129.54,129.64,128.77,130.72,128.33,128.24,129.65)
n<-30
x_bar<-mean(d8)
s<-sd(d8)
```
###(a) Can you support a claim that mean sodium content of this brand of corn flakes is 130 milligrams? Use alpha = 0.05.  h0:u=130, h1:u!=130
```{r } 
alpha <- 0.05
u0<-130
t0<-(x_bar-u0)/(s/sqrt(n))
ta<-qt(alpha/2, n-1, lower.tail=FALSE)
abs(t0)>ta
```
Since the result is false we can not reject the hypothesis, so we can support the claim that mean = 130mg.

###(b) Is there evidence that sodium content is normally distributed?
```{r }
shapiro.test(d8)
qqnorm(d8); qqline(d8)
```
Both test suggest this is not a normal distribution. 

###(c) Compute the power of the test if the true mean sodium content is 130.5 miligrams.
```{r } 
d<-(130.5-x_bar)/s
pwr.t.test(d=d,sig.level=alpha, n=n)
```
###(d) What sample size would be required to detect a true mean sodium content of 130.1 milligrams if we wanted the power of the test to be at least 0.75?
```{r } 
d<-(130.1-x_bar)/s
pwr.t.test(d=d,sig.level=alpha, power=0.75)
```
n=100.35, so 101 samples would be needed. 

###(e) Explain how the question in part (a) could be answered by constructing a two-sided confidence interval on the mean sodium content.
```{r } 
ci<-x_bar-(ta*(s/sqrt(n)))
ci2<-x_bar+(ta*(s/sqrt(n)))
```
A confidence interval is calculated to be [`r ci`, `r ci2`]. If the test mean falls within this interval, then the null hypothesis can not be rejected. Since our value falls within we can not reject the hypothesis. 

###(f) Explain how the question in part (a) could be answered by using p-value.
```{r } 
p<-2*(pt(t0, n-1, lower.tail=FALSE))
p<alpha
```
Since p<alpha is false we can not reject the hypothesis.

##Problem 9
A researcher claims that at least 10% of all football helmets have manufacturing flaws that could potentially cause injury to the wearer. A sample of 200 helmets revealed that 16 helmets contained such defects.
```{r } 
p0<-0.10
n<-200
p_bar<-16/200
```
###(a) Does this finding support the researcher's claim? Use alpha = 0.01. h0:p=0.10, h1:p<0.10
```{r } 
alpha<-0.01
z0<-(p_bar-p0)/sqrt((p0*(1-p0))/n)
za<-qnorm(1-alpha)
z0< -za
```
Since false, we can not reject the researcher's claim that at least 10% of all helmets have flaws. 

###(b) Find the P-value for this test.
```{r } 
p<-pnorm(z0, lower.tail=FALSE)
p<alpha
```
Since is false we can not reject the hypothesis. 

##Problem 10
The advertised claim for batteries for cell phones is set at 48 operating hours, with proper
charging procedures. A study of 5000 batteries is carried out and 15 stop operating prior
to 48 hours. Do these experimental results support the claim that less than 0.2 percent of
the company's batteries will fail during the advertized time period, with proper charging
procedures? Use a hypothesis-testing procedure with alpha = 0.01.
h0: p=0.002, h1:p>0.002
```{r } 
alpha<-0.01
n<-5000
x_bar<-15
p0<-0.002
p_bar<-x_bar/n
z0<-(p_bar-p0)/sqrt((p0*(1-p0))/n)
za<-qnorm(1-alpha)
z0< -za
pnorm(z0)
```
Since the test statistic is not less than the z \alpha, we fail to reject the null hypothesis. 

##Problem 11
Consider the following 75 observations.
```{r } 
d11<-c(2,3,1,0,0,0,1,1,1,0,1,1,2,1,1,2,0,0,1,1,2,4,0,1,1,2,3,1,0,0,0,0,1,2,0,0,2,2,2,2,2,1,0,0,2,0,0,1,0,0,2,3,1,2,0,4,1,1,0,2,3,1,1,0,0,2,1,3,0,2,1,1,1,0,0)
```
Based on this data, is a Poisson distribution an appropriate model?
```{r } 
freq<-as.data.frame(table(d11))
lambda<-((0*freq[1, 2])+(1*freq[2, 2])+(2*freq[3, 2])+(3*freq[4, 2])+(4*freq[5, 2]))/75
freq$efreq<-c(dpois(0,lambda)*75, dpois(1,lambda)*75, dpois(2,lambda)*75, dpois(3,lambda)*75,dpois(4,lambda)*75)
freq
k<-4
dof<-k-1-1
chi20<-(((freq$Freq[1]-freq$efreq[1])^2/freq$efreq[1])+((freq$Freq[2]-freq$efreq[2])^2/freq$efreq[2])+((freq$Freq[3]-freq$efreq[3])^2/freq$efreq[3])+(((freq$Freq[4]+freq$Freq[5])-(freq$efreq[4]+freq$efreq[4]))^2/(freq$efreq[4]+freq$efreq[4])))
chi2a<-qchisq(0.05,dof)
chi20>chi2a

```
No, since chi20 > chi2a we reject the null hypothesis that the distrubution is poisson. 

##Problem 12
Use the following 60 observations.
```{r } 
d12<-c(107.20,102.87,109.85,103.37,105.16,93.71,99.05,96.96,92.30,105.36,91.38,113.37,81.19,100.71,90.80,121.75,104.78,94.85,83.65,91.93,101.49,105.57,99.24,98.94,100.42,112.15,99.41,83.20,89.52,108.96,107.78,99.79,98.99,85.94,101.03,102.62,96.50,89.35,108.82,106.07,109.11,88.68,101.44,98.70,95.82,93.14,101.10,115.03,126.73,109.71,97.90,109.03,109.97,100.38,104.57,91.03,91.35,94.60,100.26,103.60)
n<-60
```
###Based on this data, is a normal distribution an appropriate model? To check this do the following:
###(a) Plot the normal probability plot for this data.
```{r }
qqnorm(d12); qqline(d12)
```
###(b) Perform a chi square goodness of fit test?

hist<-hist(d12, breaks=10)
k<-5
breaks_cdf <- pgamma(hist$breaks, shape=1/2, scale=(k-1-1)/2)  ##wrong shape & scale??? should be for chisquare?
probs <- rollapply(breaks_cdf, 2, function(x) x[2]-x[1])
chisq.test(hist$counts, p=probs, rescale.p=TRUE, simulate.p.value=TRUE)

###(c) Use Shapiro-Wilk test.
```{r }
shapiro.test(d12)
```

##Problem 13
Use the following 100 observations.
```{r } 
d13<-c(7.08,19.33,2.36,15.61,9.73,3.86,5.23,7.36,17.62,11.27,47.93,0.39,68.18,14.60,34.61,18.79,9.67,13.82,12.02,4.81,6.69,20.26,20.97,0.12,1.50,2.19,1.11,21.18,0.78,13.63,14.77,9.41,9.27,5.21,8.37,16.81,0.12,5.55,7.71,3.28,7.01,7.29,5.33,1.84,30.61,13.41,10.28,6.21,14.30,0.55,62.21,3.49,0.15,9.36,3.78,2.64,18.79,4.37,35.40,2.63,2.59,9.17,6.97,0.91,1.02,2.19,3.27,15.29,5.24,3.79,0.45,5.64,0.36,2.61,35.01,7.65,12.00,3.43,39.58,17.65,15.55,14.66,0.37,3.02,49.43,14.70,0.42,34.84,2.42,10.54,15.19,15.13,15.37,17.79,22.65,2.31,12.34,12.28,22.88,0.15)
n<-100
```
###(a) Use normal probability plot, chi square goodness of fit test, and Shapiro-Wilk test to check normality of data.
```{r } 
x_bar<-mean(d13)
s<-sd(d13)
alpha<-0.05
qqnorm(d13); qqline(d13)

hist<-hist(d13, breaks=10)
k<-10
breaks_cdf <- pgamma(hist$breaks, shape=1/2, scale=(k-1-1)/2)
probs <- rollapply(breaks_cdf, 2, function(x) x[2]-x[1])
chisq.test(hist$counts, p=probs, rescale.p=TRUE, simulate.p.value=TRUE)

shapiro.test(d13)
```
All three tests show this data set is not a good fit. 

###(b) Does the exponential distribution seem to be a reasonable model for these data? Perform an appropriate goodness-of-fit test to answer this question.   ?????????????
```{r } 
hist<-hist(d13, breaks=4)
k<-4
breaks_cdf <- pexp(hist$breaks, x_bar)
probs <- rollapply(breaks_cdf, 2, function(x) x[2]-x[1])
chisq.test(hist$counts, p=probs, rescale.p=TRUE, simulate.p.value=TRUE)
```